---
title: "Mini Project 1: Health Insurance Analysis"
author: "Diane Chiang"
date: "2024-01-18"
knit: (function(inputFile, encoding) {
        Sys.setenv(RSTUDIO_PANDOC="/Applications/RStudio.app/Contents/Resources/app/quarto/bin/tools");
        rmarkdown::render(inputFile,
                          encoding=encoding, 
                          output_file=file.path(dirname(inputFile), "hello.pdf")) })
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
HealthInsurance <- read.csv("health_insurance.csv")
head(HealthInsurance)
```

## Part A
### 1. What proportion of the individuals in this dataset have health insurance?
```{r}
total_p = nrow(HealthInsurance)
p_hi = nrow(HealthInsurance[HealthInsurance["insurance"] == "yes",])
prop_p_hi = p_hi / total_p
print(prop_p_hi)
```
About 80.12% of the individuals has health insurance in this dataset.


### 2. Construct a confidence interval for the true population proportion of individuals who have health insurance:
#### a. What is the estimate for the standard error for the sample proportion of individuals who have health insurance?
```{r}
# standard error = sqrt((p * q)/n)
std_error = sqrt((prop_p_hi * (1 - prop_p_hi)) / total_p)
print(std_error)
```
The estimate for standard error of the sample proportions is around 0.00425

#### b. Construct an (approximate) 95% confidence interval for the true population proportion of individuals who have health insurance
```{r}
# conf_int = p' -+ 1.96 * standard_error
lower_b = prop_p_hi - 1.96 * std_error
upper_b = prop_p_hi + 1.96 * std_error

print(c(lower_b, upper_b))
```

#### c. Provide an interpretation for this confidence interval.
We estimate with 95% confidence that the true population proportion of individuals who have health insurance is approximately between 0.7928 and 0.8095.

### 3. According to census.gov (https://www.census.gov/library/publications/1996/demo/p60-195.html), an estimated 15.4% of individuals were without health insurance in 1995. Taking this 15.4% as the truth and using the data collected from this survey, is there evidence that the proportion of individuals with health insurance changed from 1995 to 1996?
#### a. Set up the null and alternative hypothesis for this test. Define any relevant notation.
Let p be the proportion of individuals with health insurance in 1996. The null hypothesis is that p = 0.846 and the alternate hypothesis is that p != 0.846.

#### b. Which sort of test would be appropriate to use here?
I would conduct the two-tailed significance test of a single population proportion.

#### c. Calculate the relevant test statistic.
```{r}
# prop_p_wo_hi = 1 - prop_p_hi
sd = sqrt((0.846 * (1 - 0.846)) / total_p)
z_score = (prop_p_hi - 0.846) / sd
print(z_score)
print(2*pnorm(q=z_score, lower.tail=TRUE))

test = prop.test(x = p_hi, n = total_p, p = 0.846, alternative = "two.sided", correct = FALSE)
test
```